---
title: "Case study Project 1"
author: "Ly Le Thi"
date: "2023-10-27"
output:
  pdf_document: {}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load library


```{r}
rm(list = ls())
```

```{r, warning=FALSE}
library(tidyverse)
library(lubridate)
library(readxl)
library(dplyr) 
library(ggplot2)
library(zoo)
library(here)
library(readxl)
library(dplyr)
library(magrittr)
library(ggplot2)
library(gridExtra)
library(mice)
library(quantmod)
library(tsibble)
library(stats)
library(purrr)
library(rlang)
library(forecast)
library(lmSubsets)
```

## a. Import dataset

```{r}
tsdata <- read_excel("~/Desktop/case study/PredictorData2022.xlsx")
ts_quarter <- read_excel("~/Desktop/case study/PredictorData2022.xlsx", sheet = "Quarterly")
ts_year <- read_excel("~/Desktop/case study/PredictorData2022.xlsx", sheet = "Annual")


tsdata$yyyymm <- as.yearmon(as.character(tsdata$yyyymm), format="%Y%m")
ts_quarter$yyyyq <- as.yearqtr(as.character(ts_quarter$yyyyq), format="%Y%q")
ts_year$yyyy <- as.Date(as.character(ts_year$yyyy), format="%Y")

tsdata %<>% select(-CRSP_SPvw, -CRSP_SPvwx, -csp)
ts_quarter %<>% select(-CRSP_SPvw, -CRSP_SPvwx, -csp)
ts_year %<>% select(-CRSP_SPvw, -CRSP_SPvwx, -csp)
```

```{r}
vars_monthly <- names(tsdata)
vars_quarterly <-names(ts_quarter) 
vars_annual <-names(ts_year)

setdiff(vars_quarterly,vars_monthly)
setdiff(vars_annual,vars_monthly)
```

```{r}
# Rename the column
colname_mapping <- c(
"time" = "yyyymm",
  "time" = "yyyyq",
  "time" = "yyyy",
  "sp_index" = "Index",
  "dividend_12" = "D12",
  "earning_12" = "E12",
  "bookmarketRatio" = "b/m",
  "treasury_bill" = "tbl",
  "cor_AAA" = "AAA",
  "cor_BAA" = "BAA",
  "longterm_yeild" = "lty",
  "net_equity" = "ntis",
  "risk_free" = "Rfree",
  "inflation" = "infl",
  "long_returnrate" = "ltr",
  "cbond_return" = "corpr",
  "stock_var" = "svar",
  "cs_premium" = "csp",
  "consumWealIncome" = "cay",
  "investmentCapital" = "ik",
  "dividend3year" = "D3",
  "earning3year" = "E3",
  "consum_wealth_monthly" = "caym",
  "equity_issuing" = "eqis"
)

#Rename columns 
filtered_dict_monthly <- colname_mapping[colname_mapping %in% colnames(tsdata)]
tsdata <- rename(tsdata, !!!filtered_dict_monthly)

filtered_dict_quarterly <- colname_mapping[colname_mapping %in% colnames(ts_quarter)]
ts_quarter <- rename(ts_quarter, !!!filtered_dict_quarterly)

filtered_dict_annual <- colname_mapping[colname_mapping %in% colnames(ts_year)]
ts_year <- rename(ts_year, !!!filtered_dict_annual)

```

# Format dataframe
```{r}
# Transform variables into numeric
for(i in names(tsdata)) {
  if(class(tsdata[[i]]) == "character") {
    tsdata[[i]] <- as.numeric(tsdata[[i]])
  }
}

for(i in names(ts_quarter)) {
  if(class(ts_quarter[[i]]) == "character") {
    ts_quarter[[i]] <- as.numeric(ts_quarter[[i]])
  }
}

for(i in names(ts_year)) {
  if(class(ts_year[[i]]) == "character") {
    ts_year[[i]] <- as.numeric(ts_year[[i]])
  }
}
```


## b. Generate the excess returns series

```{r}
calculate_returns <- function(input){
  input %<>% mutate(returns = as.vector(quantmod::Delt(sp_index))) 
  input %<>% mutate(excess_returns = returns - risk_free)
  return(input)
}
```

```{r}
# Create subsets for different frequencies --------------------------------
tsdata %<>% calculate_returns()
ts_quarter %<>% calculate_returns()
ts_year %<>% calculate_returns()
```

```{r}

```


## c. Structure of excess return series

# Plot excess return series

```{r, warning=FALSE}
excess_returns_plot <- ggplot(tsdata, aes(x = time, y = excess_returns)) +
  geom_line() +
  labs(title = "Time Series Plot of Monthly Excess Returns", x = "Time", y = "Excess Return")+
  theme_bw()
print(excess_returns_plot)
```

```{r, warning=FALSE}
excess_returns_plotq <- ggplot(ts_quarter, aes(x = time, y = excess_returns)) +
  geom_line() +
  labs(title = "Time Series Plot of Quaterly Excess Returns", x = "Time", y = "Excess Return")+
  theme_bw()

print(excess_returns_plotq)
```

```{r, warning=FALSE}
excess_returns_ploty <- ggplot(ts_year, aes(x = time, y = excess_returns)) +
  geom_line() +
  labs(title = "Time Series Plot of Yearly Excess Returns", x = "Time", y = "Excess Return")+
  theme_bw()

print(excess_returns_ploty)

```
```{r}
summary(tsdata$excess_returns)
summary(ts_quarter$excess_returns)
summary(ts_year$excess_returns)
tsdata$excess_returns %>% var(na.rm=TRUE)
ts_quarter$excess_returns %>% var(na.rm=TRUE)
ts_year$excess_returns %>% var(na.rm=TRUE)
```

## Plot acf and pacf
# Monthly

```{r}
excess_rt <- na.omit(tsdata$excess_returns)
```

```{r}
acf_result <- Acf(excess_rt, lag.max = 60, main = "ACF of Excess Return Monthly")
```

```{r}
pacf_result <- pacf(excess_rt,lag.max = 60, main = "PACF of Excess Return Mothly")
```
Look more likely AR(1) model

# Quaterly

```{r}
excess_rtq <- na.omit(ts_quarter$excess_returns)
```

```{r}
acf_resultq <- Acf(excess_rtq, lag.max = 60, main = "ACF of Excess Return Quarterly")
```

```{r}
pacf_resultq <- pacf(excess_rtq,lag.max = 60, main = "PACF of Excess Return Quarterly")
```
Look like AR(3, AR(4) 

# Yearly

```{r}
excess_rty <- na.omit(ts_year$excess_returns)
```

```{r}
acf_resulty <- Acf(excess_rty, main = "ACF of Excess Return Annual")
```

```{r}
pacf_resulty <- pacf(excess_rty, main = "PACF of Excess Return Annual")
```
Look like no correlation. possibly can fit AR(2)

## d.e. Fit an AR models base on Information criteria and generate forecast

# For monthly data

```{r}
ar_modelling <- function(input) {
  data_cleaned <- input %>% filter(!is.na(excess_returns))
  
  p_values <- 1:2
  aic_values <- numeric(length(p_values))
  ar_models <- list()
  
  for (p in p_values) {
    AR_model <- arima(data_cleaned$excess_returns, order = c(p, 0, 0), method = "ML")
    
    aic_values[p] <- AIC(AR_model)
    
    ar_models[[p]] <- AR_model
  }
  
  best_p <- p_values[which.min(aic_values)]
  best_AR <- ar_models[[best_p]]
  
  # Calculate fitted values
  fitted_values <- data_cleaned$excess_returns - best_AR$residuals
  
  plot_data <- data.frame(
    Date = as.Date(input$time[!is.na(input$excess_returns)]),
    ExcessReturns = data_cleaned$excess_returns,
    Fitted = fitted_values
  )
  
  MSFE <- sqrt(mean(best_AR$residuals^2))
  
  # Plot data
  plot <- ggplot(plot_data, aes(x = Date)) +
    geom_line(aes(y = ExcessReturns), color = "black") +
    geom_line(aes(y = Fitted), color = "red") +
    labs(title = paste("Time Series with AR(", best_p, ") Fitted Values")) +
    theme_bw()
  
  result <- list(best_p = best_p, aic_values = aic_values, AR_model = best_AR, MSFE = MSFE, plot = plot)
  
  return(result)
}

```

```{r, warning=FALSE}
result <- ar_modelling(input = tsdatatrain)
print(paste("Best AR(p) order:", result$best_p))
summary(result$AR_model)
print(paste("AIC values for p =", 1:2, ":", result$aic_values))
print(paste("MSFE:", result$MSFE))
print(result$plot)
```

```{r}
ar_modelling <- function(input) {
  data_cleaned <- input %>% filter(!is.na(excess_returns))
  
  p_values <- 1:5
  aic_values <- numeric(length(p_values))
  ar_models <- list()
  
  for (p in p_values) {
    AR_model <- arima(data_cleaned$excess_returns, order = c(p, 0, 0), method = "ML")
    
    aic_values[p] <- AIC(AR_model)
    
    ar_models[[p]] <- AR_model
  }
  
  best_p <- p_values[which.min(aic_values)]
  best_AR <- ar_models[[best_p]]
  
  # Calculate fitted values
  fitted_values <- data_cleaned$excess_returns - best_AR$residuals
  
  plot_data <- data.frame(
    Date = as.Date(input$time[!is.na(input$excess_returns)]),
    ExcessReturns = data_cleaned$excess_returns,
    Fitted = fitted_values
  )
  
  MSFE <- sqrt(mean(best_AR$residuals^2))
  
  # Plot data
  plot <- ggplot(plot_data, aes(x = Date)) +
    geom_line(aes(y = ExcessReturns), color = "black") +
    geom_line(aes(y = Fitted), color = "red") +
    labs(title = paste("Time Series with AR(", best_p, ") Fitted Values")) +
    theme_bw()
  
  result <- list(best_p = best_p, aic_values = aic_values, AR_model = best_AR, MSFE = MSFE, plot = plot)
  
  return(result)
}
```



```{r}
result <- ar_modelling(input = ts_quarter)
print(paste("Best AR(p) order:", result$best_p))
print(paste("AIC values for p =", 1:5, ":", result$aic_values))
summary(result$AR_model)
print(paste("MSFE:", result$MSFE))
print(result$plot)
```


```{r}
result <- ar_modelling(input = ts_year)
print(paste("Best AR(p) order:", result$best_p))
print(paste("AIC values for p =", 1:5, ":", result$aic_values))
summary(result$AR_model)
print(paste("MSFE:", result$MSFE))
print(result$plot)
```



## f. Linear predictive model

```{r}
regressors_all <- union(union(names(ts_year), names(tsdata)), names(ts_quarter)) %>% setdiff(c("risk_free", "sp_index", "returns", "excess_returns", "time"))
regressors_monthly <- names(tsdata) %>% setdiff(c( "risk_free","sp_index", "returns", "excess_returns", "time"))
regressors_quarterly <- names(ts_quarter) %>% setdiff(c( "risk_free","sp_index", "returns", "excess_returns","time"))
regressors_annual <- names(ts_year) %>% setdiff(c("risk_free","sp_index", "returns", "excess_returns", "time"))

```


```{r}
lagged_regressors_all <- paste0("lag(", regressors_all, ")")
lagged_regressors_monthly <- paste0("lag(", regressors_monthly, ")")
lagged_regressors_quarterly <- paste0("lag(", regressors_quarterly, ")")
lagged_regressors_annual <- paste0("lag(", regressors_annual, ")")

MSFE_df <- data.frame(monthly=rep(NA,19),
                      quarterly=rep(NA,19),
                      annual=rep(NA,19),
                      row.names = c(regressors_all,"all predictors", "AR(p)"))

datasets <- list(tsdata, ts_quarter, ts_year)

```



# Models monthly

```{r, warning=FALSE}
# Create an empty data frame to store MSFE values
MSFE_df_monthly <- data.frame(Model = c(lagged_regressors_monthly, "all predictors", "AR(p)"), 
                              MSFE = rep(NA, length(lagged_regressors_monthly) + 2))

for (i in 1:length(MSFE_df_monthly$Model)) {
  if (MSFE_df_monthly$Model[i] == "AR(p)") {
    # Try to calculate MSFE for AR(p) model
    tryCatch({
      ar_result <- ar_modelling(input = tsdata)
      MSFE_df_monthly$MSFE[i] <- ar_result$MSFE_aic
    }, error = function(e) {
      cat("Warning: AR(p) model did not converge or fit successfully.\n")
    })
  } else if (MSFE_df_monthly$Model[i] == "all predictors") {
    # LM for all predictors
    formula_all_predictors1 <- paste("excess_returns ~", paste(lagged_regressors_monthly, collapse = " + ")) %>% as.formula()
    model_all_predictors <- lm(formula_all_predictors1, data = tsdata)
    MSFE_df_monthly$MSFE[i] <- mean(model_all_predictors$residuals^2)^0.5
  } else {
    # LM for single predictor
    formula <- paste("excess_returns ~ ", MSFE_df_monthly$Model[i], sep = "") %>% as.formula()
    model <- lm(formula, data = tsdata)
    MSFE_df_monthly$MSFE[i] <- mean(model$residuals^2)^0.5
  }
}

# Print or return the MSFE data frame
print(MSFE_df_monthly)

```


```{r, warning=FALSE}
# Create an empty data frame to store MSFE values
MSFE_df_quarterly <- data.frame(Model = c(lagged_regressors_quarterly, "all predictors", "AR(p)"), MSFE = rep(NA, length(lagged_regressors_quarterly) + 2))

for (i in 1:length(MSFE_df_quarterly$Model)) {
  if (MSFE_df_quarterly$Model[i] == "AR(p)") {
    # Try to calculate MSFE for AR(p) model
    tryCatch({
      ar_result <- ar_modelling(input = ts_quarter)
      MSFE_df_quarterly$MSFE[i] <- ar_result$MSFE_aic
    }, error = function(e) {
      cat("Warning: AR(p) model did not converge or fit successfully.\n")
    })
  } else if (MSFE_df_quarterly$Model[i] == "all predictors") {
    # LM for all predictors
    formula_all_predictors <- paste("excess_returns ~", paste(lagged_regressors_quarterly, collapse = " + ")) %>% as.formula()
    model_all_predictors <- lm(formula_all_predictors, data = ts_quarter)
    MSFE_df_quarterly$MSFE[i] <- mean(model_all_predictors$residuals^2)^0.5
  } else {
    # LM for single predictor
    formula <- paste("excess_returns ~ ", MSFE_df_quarterly$Model[i], sep = "") %>% as.formula()
    model <- lm(formula, data = ts_quarter)
    MSFE_df_quarterly$MSFE[i] <- mean(model$residuals^2)^0.5
  }
}

# Print or return the MSFE data frame
print(MSFE_df_quarterly)

```

```{r}
MSFE_df_annual <- data.frame(Model = c(lagged_regressors_annual, "all predictors", "AR(p)"), 
                             MSFE = rep(NA, length(lagged_regressors_annual) + 2))


#annual data
for (i in 1:length(MSFE_df_annual$Model)) {
  if (MSFE_df_annual$Model[i] == "AR(p)") {
    # Try to calculate MSFE for AR(p) model
    tryCatch({
      ar_result <- ar_modelling(input = ts_year)
      MSFE_df_annual$MSFE[i] <- ar_result$MSFE_aic
    }, error = function(e) {
      cat("Warning: AR(p) model did not converge or fit successfully.\n")
    })
  } else if (MSFE_df_annual$Model[i] == "all predictors") {
    # LM for all predictors
    formula_all_predictors <- paste("excess_returns ~", paste(lagged_regressors_annual, collapse = " + ")) %>% as.formula()
    model_all_predictors <- lm(formula_all_predictors, data = ts_year)
    MSFE_df_annual$MSFE[i] <- mean(model_all_predictors$residuals^2)^0.5
  } else {
    # LM for single predictor
    formula <- paste("excess_returns ~ ", MSFE_df_annual$Model[i], sep = "") %>% as.formula()
    model <- lm(formula, data = ts_year)
    MSFE_df_annual$MSFE[i] <- mean(model$residuals^2)^0.5
  }
}

# Print or return the MSFE data frame
print(MSFE_df_annual)
```


```{r}
merged_12 <- merge(MSFE_df_monthly, MSFE_df_quarterly, by="row.names", all=TRUE)
rownames(merged_12) <- merged_12$Row.names
merged_12 %<>% select(-Row.names)
final_merged <- merge(merged_12, MSFE_df_annual, by="row.names", all=TRUE)
rownames(final_merged) <- final_merged$Row.names
final_merged %<>% select(-Row.names)
final_merged
```


```{r}
# Minimum MSFE
final_merged %>% select(MSFE.x) %>% filter(MSFE.x == min(MSFE.x, na.rm = TRUE))
final_merged %>% select(MSFE.y) %>% filter(MSFE.y == min(MSFE.y, na.rm = TRUE))
final_merged %>% select(MSFE) %>% filter(MSFE == min(MSFE, na.rm = TRUE))
```


```{r}
# Maximum MSFE
final_merged %>% select(MSFE.x) %>% filter(MSFE.x == max(MSFE.x, na.rm = TRUE))
final_merged %>% select(MSFE.y) %>% filter(MSFE.y == max(MSFE.y, na.rm = TRUE))
final_merged %>% select(MSFE) %>% filter(MSFE == max(MSFE, na.rm = TRUE))
```




## h. select predictors


```{r}
formula_all_predictors_monthly <- paste("excess_returns ~", paste(lagged_regressors_monthly, collapse=" + ")) %>% as.formula()
best_model_aic <- lm(lmSubsets::lmSelect(formula_all_predictors_monthly, data=tsdata, penalty ="AIC") %>% formula(best=1), data=tsdata)
summary(best_model_aic)
mean(best_model_aic$residuals^2)^0.5
```


```{r}
subset_quarterly <- na.omit(ts_quarter)

formula_all_predictors_quarterly <- paste("excess_returns ~", paste(lagged_regressors_quarterly, collapse=" + ")) %>% as.formula()
best_model_aic <- lm(lmSubsets::lmSelect(formula_all_predictors_quarterly, data=subset_quarterly, penalty ="AIC") %>% formula(best=1), data=subset_quarterly)
summary(best_model_aic)
mean(best_model_aic$residuals^2)^0.5

```


```{r}

subset_year <- na.omit(ts_year)

formula_all_predictors_annual <- paste("excess_returns ~", paste(lagged_regressors_annual, collapse=" + ")) %>% as.formula()
best_model_aic <- lm(lmSubsets::lmSelect(formula_all_predictors_annual, data=subset_year, penalty ="AIC") %>% formula(best=1), data=subset_year)
summary(best_model_aic)
mean(best_model_aic$residuals^2)^0.5

```


## i. Forward stepwise model selection

```{r}
lag_predictors <- function(input){
  output <- input %>%
    mutate(across(-c(time, excess_returns,risk_free, sp_index,returns), lag, .names = "lag_{.col}")) %>% 
    select(time, excess_returns, starts_with("lag_")) %>% 
    slice(-1) 
  return(output)
}
```

```{r}
lag_month<- lag_predictors(tsdata)
lag_quarter <- lag_predictors(ts_quarter)
lag_year <- lag_predictors(ts_year)
```


```{r}
get_time_span <- function(data_input, model_input){
  coefficients_names <- model_input %>% coef() %>% names() %>% setdiff("(Intercept)")
  complete_cases <- which(complete.cases(data_input[c("excess_returns", coefficients_names)]))
  used_rows <- row.names(model_input$model)
  removed_rows <- setdiff(complete_cases, as.numeric(used_rows))
  included_time_spans <- data_input$time[as.numeric(used_rows)]
  output <- paste(included_time_spans %>% min() %>% year() %>% as.character(), "-", included_time_spans %>% max() %>% year() %>% as.character())
  return(output)
}
```



```{r}
stepwise_function <- function(input){
  input_name <- deparse(substitute(input))
  
  model <- step(object= lm(excess_returns ~ 1, data = input %>% na.omit()) 
                , scope = paste("excess_returns ~", paste(c(names(input %>% select(-excess_returns, -time))), collapse=" + ")) %>% as.formula()
                , direction = "forward", trace = FALSE, k=2) 
  
  output <- data.frame(
    Data = get_time_span(data_input=input, model_input=model),
    MSFE = mean(model$residuals^2)^0.5 ,
    R2 = summary(model)$r.squared
  )
  
  
  print(summary(model))
  return(output)
}
```



```{r}
stepwise_monthly <- stepwise_function(input=lag_month)
stepwise_quarterly <- stepwise_function(input=lag_quarter)
stepwise_annual <- stepwise_function(input=lag_year)
colnames(stepwise_monthly) <- c("Monthly","MSFE_monthly", "R2_monthly")
colnames(stepwise_quarterly) <- c("Quarterly","MSFE_quarterly", "R2_quarterly")
colnames(stepwise_annual) <- c("Annual","MSFE_annual", "R2_annual")
stepwise_complete <- cbind(stepwise_monthly, stepwise_quarterly, stepwise_annual)
rownames(stepwise_complete) <- c("stepwise LM")

```

```{r}
stepwise_complete
```














